---
title: "US Census Demographic Data"
author: "Data lovers"
date: "06 Dec 2021"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```


### Chapter 1:Introduction.

Census is a very important matter, not just at the government level, but for individuals as well. The census population count decides how many congressional members each state will have for the following ten years, as well as how much federal funds localities will receive for roads, schools, housing, and social programs. 

Poverty is a complicated issue. There are numerous elements to it, as well as numerous reasons. The most common definition of poverty focuses on financial hardship. This definition defines poverty rates for towns and countries based on income inequality and financially established poverty lines and quantifies poverty by the amount of money a person makes.

Poverty becomes relative in this way. In the United States, the poverty line for a family of four is little over $26,000 a year. That amount of money is more than 36 times what a family of four living in absolute poverty in a low-income country is forced to subsist on each year.

Many families in the world's poorest countries must make do with less than $1.90 per day to cover their basic requirements. This sort of extreme poverty affects nearly 10% of the world's population.

Understanding poverty solely in terms of economic disparity and calculating relative poverty thresholds makes it impossible to see poverty as anything other than unemployment, bad living conditions, and a low income. Poverty, on the other hand, is a much broader concept.

A fair definition of poverty must take into account the many various aspects of poverty, such as hunger and a lack of shelter, illiteracy and lack of access to education, being sick and unable to see a doctor, and fearing for one's future.

By exploring this data and analyzing it we can come up with many insights, such as how the population varies across the states, which states that have the best transportation, how do income and poverty look like across US, employment/unemployment rates and so many questions that can be answered from these data. 

Therefore, our aim in this project is to analyze the US census of 2015 to answer our interesting question: 


1. Does a relationship exist between ‘child poverty’ and ‘unemployment’? Or how can we predict the unemployment rate?


2. Does the data support ‘professional’ (individual who has a professional job) influence ‘income’?


3. How can we predict the average income in counties? 



### Chapter 2:Description of the Data.

•	Dataset that is used for our project is “US Census Demographic Data” taken from the common data source Kaggle (which is an online community of data scientists and machine learning practitioners.

Link: https://www.kaggle.com/muonneutrino/us-census-demographic-data?select=acs2015_county_data.csv

•	Dataset contains information about (US Census 2015) American Community Survey 5-year estimates. 

•	Number of variables in the dataset is 37.

•	Number of observations in the dataset is 3220.

•	To perform Exploratory Data Analysis, we primarily focus on the following aspects(variables) from the dataset:

* TotalPop (Total population for county).

*	Income. (Income in county).

*	Poverty (Poverty rate in county).

*	ChildPoverty. (Child poverty rate in county).

* Employed (Total employed for county).

*	Unemployment. (Unemployment rate for county).

* Production. (Rate of this job sector in county).

* WorkAtHome. (Work-at-home rate in county).

* PublicWork. (Public work rate in county).

* SelfEmployed. (Self-employed rate in county).





**This is our dataset before performing any cleansing or dropping columns: **

```{r step1, results='hold'}


# importing the dataset as Adata, and making sure the data type for the variables are set properly for categorical variables.**

Adata1 <- data.frame(read.csv("acs2015_county_data.csv",header = TRUE))

str(Adata1)

#summary(Adata1)

#  we don't need to change any variables type, all of them set properly.
# There are a lot of variables , we may drop some of them, like :
#Hispanic, White,	Black,	Native,	Asian	Pacific, IncomeErr,	IncomePerCap,	IncomePerCapErr


# Dropping columns as agreed.  

Adata <- data.frame (subset(Adata1, select = -c(Hispanic,White,Black, Native, Asian, Pacific,IncomeErr, IncomePerCap, IncomePerCapErr, Drive, Carpool, Transit, Walk, OtherTransp, MeanCommute)))

str(Adata)
#summary(Adata)
  

```



**And here is the summary of our dataset **
```{r summary, results='hold'}


summary(Adata)


```


### Chapter 3: Data Cleaning.


As we did in the previous project, we performed data cleansing and normality tests to prepare our data for EDA. 

we performed the following steps:

1- Drop the unnecessary columns. 

2- Remove the outliers. 

3- Remove the Null values if exist.


Hint:To detect the outliers, we first polt it using Hitogram , then remove outliers, finally polt it again to check after removing the outliers. Same technique applied for all the 10 variables. 


```{r step2, include=FALSE}

# After reading the data we need now to clean it to prepare for the analysis. 
# We started with "income", first we polte it to identify the outlier and then removing it. 
# Also , we removed the null values because it make error while do mathematics calculation. 
# Then we follow the same steps for the remaining columns that we will gonna use in this phase. 
# Currently our desired columns: TotalPop, Income, Poverty, ChildPoverty, Employed, Unemployment, Production, WorkAtHome,PublicWork, SelfEmployed

################################################################################


# First, lets take a look at the histogram for the income to detect the outliers 



hist(Adata$Income,
  xlab = "Income",
  main = "Histogram of Income",
  breaks = sqrt(nrow(Adata))
) # set number of bins




# # Now, we can extract the outliers from the incom using boxplot.stats()$out function.  
 income_out <- boxplot.stats(Adata$Income)$out
 income_out


# To count how many outliers in the income
 x <- data.frame(income_out)
x 
 nrow(x) # outliers numbers for income


 #Now, we will remove the outliers from our dataset and store the clean result in Adata2

  Adata2<- Adata[-which(Adata$Income %in% income_out),] # here we removed income outlier


 Adata2
 
 hist(Adata2$Income,
  xlab = "Income",
  main = "Histogram of Income",
  breaks = sqrt(nrow(Adata2))
) # set number of bins

 
# #Adata
# Adata2 


 # Now we will remove the naull values from the incmoe to allow for math calculations. 
 
 my.null <- data.frame (subset(Adata2,subset= is.na(Adata2$Income)))# To indicate the null values of income. 
my.null


Adata3 <- data.frame (subset(Adata2, subset= CensusId!= 48301 )) # To remove the record that contains null value

# Now we can calculate the max, mean , min .. etc of the income
max(Adata3$Income)
mean(Adata3$Income)
min(Adata3$Income)



```



```{r step3, include=FALSE}

# same thing for population column. clean the data. 

# First, lets take a look at the histogram for the population to detect the outliers 



hist(Adata3$TotalPop,
  xlab = "TotalPop",
  main = "Histogram of Population",
  breaks = sqrt(nrow(Adata3))
) # set number of bins




# # Now, we can extract the outliers from the population using boxplot.stats()$out function.  
 pop_out <- boxplot.stats(Adata3$TotalPop)$out
 pop_out


# To count how many outliers in the population
 x2 <- data.frame(pop_out)
x2 
 nrow(x2) # outliers numbers 


 #Now, we will remove the outliers from our dataset and store the clean result in Adata4

  Adata4<- Adata3[-which(Adata3$TotalPop %in% pop_out),] # here we removed  outlier


 Adata4
 
 hist(Adata4$TotalPop,
  xlab = "TotalPop",
  main = "Histogram of Population",
  breaks = sqrt(nrow(Adata4))
) # set number of bins

 
# #Adata
# Adata2 


 # Now we will remove the naull values from the population to allow for math calculations. 
 
 my.null2 <- data.frame (subset(Adata4,subset= is.na(Adata4$TotalPop)))# To indicate the null values of Population 
my.null2  # No null values. 


# Now we can calculate the max and the mean .. etc of the population. 
max(Adata4$TotalPop)
mean(Adata4$TotalPop)
min(Adata4$TotalPop)



```


```{r step4, include=FALSE}


# same thing for poverty column. clean the data. 

# First, lets take a look at the histogram for the poverty to detect the outliers 



hist(Adata4$Poverty,
  xlab = "Poverty",
  main = "Histogram of Poverty",
  breaks = sqrt(nrow(Adata4))
) # set number of bins




# # Now, we can extract the outliers from the population using boxplot.stats()$out function.  
 Poverty_out <- boxplot.stats(Adata4$Poverty)$out
 Poverty_out


# To count how many outliers in the population
 x3 <- data.frame(Poverty_out)
x3 
 nrow(x3) # outliers numbers 


 #Now, we will remove the outliers from our dataset and store the clean result in Adata5

  Adata5<- Adata4[-which(Adata4$Poverty %in% Poverty_out),] # here we removed  outlier


 Adata5
 
 hist(Adata5$Poverty,
  xlab = "Poverty",
  main = "Histogram of Poverty",
  breaks = sqrt(nrow(Adata5))
) # set number of bins

 
 



 # No null values. 
 

# Now we can calculate the max and the mean .. etc of the population. 
max(Adata4$Poverty)
mean(Adata4$Poverty)
min (Adata4$Poverty)






```



```{r step5, include=FALSE}


# same thing for Child Poverty column. clean the data. 

# First, lets take a look at the histogram for the Child Poverty to detect the outliers 



hist(Adata5$ChildPoverty,
  xlab = "ChildPoverty",
  main = "Histogram of ChildPoverty",
  breaks = sqrt(nrow(Adata5))
) # set number of bins




# # Now, we can extract the outliers using boxplot.stats()$out function.  
 ChildPoverty_out <- boxplot.stats(Adata5$ChildPoverty)$out
 ChildPoverty_out


# To count how many outliers in the population
 x4 <- data.frame(ChildPoverty_out)
x4
 nrow(x4) # outliers numbers 


 #Now, we will remove the outliers from our dataset and store the clean result in Adata6

  Adata6<- Adata5[-which(Adata5$ChildPoverty %in% ChildPoverty_out),] # here we removed  outlier


 Adata6
 
 hist(Adata6$ChildPoverty,
  xlab = "ChildPoverty",
  main = "Histogram of Child Poverty",
  breaks = sqrt(nrow(Adata6))
) # set number of bins

 
 

 
 my.null3 <- data.frame (subset(Adata6,subset= is.na(Adata6$ChildPoverty)))# To indicate the null values
my.null3  
 

 Adata7 <- data.frame (subset(Adata6, subset= CensusId!= 15005 )) # To remove the record that contains null value



# Now we can calculate the max and the mean .. etc of the population. 
max(Adata7$ChildPoverty)
mean(Adata7$ChildPoverty)
min (Adata7$ChildPoverty)






```




```{r step6, include=FALSE}

# same thing for Employed column. clean the data. 

# First, lets take a look at the histogram for the Employed to detect the outliers 



hist(Adata7$Employed,
  xlab = "Employed",
  main = "Histogram of Employed",
  breaks = sqrt(nrow(Adata7))
) # set number of bins




# # Now, we can extract the outliers using boxplot.stats()$out function.  
 Employed_out <- boxplot.stats(Adata7$Employed)$out
 Employed_out


# To count how many outliers in the population
 x5 <- data.frame(Employed_out)
x5
 nrow(x5) # outliers numbers 


 #Now, we will remove the outliers from our dataset and store the clean result in Adata8

  Adata8<- Adata7[-which(Adata7$Employed %in% Employed_out),] # here we removed outlier


 Adata8
 
 hist(Adata8$Employed,
  xlab = "Employed",
  main = "Histogram of Employed",
  breaks = sqrt(nrow(Adata8))
) # set number of bins

 
 
# No null values. 
 

# # Now we can calculate the max and the mean .. etc of the population. 
max(Adata8$Employed)
mean(Adata8$Employed)
min (Adata8$Employed)



```




```{r step7, include=FALSE}


# same thing for Unemployment column. clean the data. 

# First, lets take a look at the histogram for the Unemployment to detect the outliers 



hist(Adata8$Unemployment,
  xlab = "Unemployment",
  main = "Histogram of Unemployment",
  breaks = sqrt(nrow(Adata8))
) # set number of bins




 # Now, we can extract the outliers using boxplot.stats()$out function.  
 Unemployment_out <- boxplot.stats(Adata8$Unemployment)$out
 Unemployment_out
# 
# 
#  To count how many outliers in the population
  x6 <- data.frame(Unemployment_out)
 x6
  nrow(x6) # outliers numbers 
# 
# 
 #Now, we will remove the outliers from our dataset and store the clean result in Adata9
# 
   Adata10<- Adata8[-which(Adata8$Unemploymen %in% Unemployment_out),] # here we removed outlier
# 
# 
 Adata10
#  
  hist(Adata10$Unemployment,
   xlab = "Unemployment",
   main = "Histogram of Unemployment",
   breaks = sqrt(nrow(Adata10))
) # set number of bins
  


# # No null values. 
 
# 
# # # Now we can calculate the max and the mean .. etc . 
 max(Adata10$Unemployment)
 mean(Adata10$Unemployment)
min (Adata10$Unemployment)
 



```




```{r step 8, include=FALSE}


# same thing for Production column. clean the data. 

# First, lets take a look at the histogram for the Production to detect the outliers 



hist(Adata10$Production,
  xlab = "Production",
  main = "Histogram of Production",
  breaks = sqrt(nrow(Adata10))
) # set number of bins




 # Now, we can extract the outliers using boxplot.stats()$out function.  
 production_out <- boxplot.stats(Adata10$Production)$out
 production_out
# 
# 
#  To count how many outliers in the population
  x7 <- data.frame(production_out)
 x7
  nrow(x7) # outliers numbers 
# 
# 
 #Now, we will remove the outliers from our dataset and store the clean result in Adata11
# 
   Adata11<- Adata10[-which(Adata10$Production %in% production_out),] # here we removed outlier
# 
# 
 Adata11
#  
  hist(Adata11$Production,
   xlab = "Production",
   main = "Histogram of Production",
   breaks = sqrt(nrow(Adata11))
) # set number of bins
  


# # No null values. 
 
# 
# # # Now we can calculate the max and the mean .. etc . 
 max(Adata11$Production)
 mean(Adata11$Production)
min (Adata11$Production)
 






```

 

```{r step 9, include=FALSE}


# same thing for WorkAtHome column. clean the data. 

# First, lets take a look at the histogram for the WorkAtHome to detect the outliers 



hist(Adata11$WorkAtHome,
  xlab = "WorkAtHome",
  main = "Histogram of WorkAtHome",
  breaks = sqrt(nrow(Adata11))
) # set number of bins




 # Now, we can extract the outliers using boxplot.stats()$out function.  
 WorkAtHome_out <- boxplot.stats(Adata11$WorkAtHome)$out
 WorkAtHome_out
# 
# 
#  To count how many outliers in the population
  x8 <- data.frame(WorkAtHome_out)
 x8
  nrow(x8) # outliers numbers 
# 
# 
 #Now, we will remove the outliers from our dataset and store the clean result in Adata11
# 
   Adata12<- Adata11[-which(Adata11$WorkAtHome %in% WorkAtHome_out),] # here we removed outlier
# 
# 
 Adata12
#  
  hist(Adata12$WorkAtHome,
   xlab = "WorkAtHome",
   main = "Histogram of WorkAtHome",
   breaks = sqrt(nrow(Adata12))
) # set number of bins
  


# # No null values. 
 
# 
# # # Now we can calculate the max and the mean .. etc . 
 max(Adata12$WorkAtHome)
 mean(Adata12$WorkAtHome)
min (Adata12$WorkAtHome)
 






```




```{r step10, include=FALSE}


# same thing for PublicWork column. clean the data. 

# First, lets take a look at the histogram for the PublicWork to detect the outliers 



hist(Adata12$PublicWork,
  xlab = "PublicWork",
  main = "Histogram of PublicWork",
  breaks = sqrt(nrow(Adata12))
) # set number of bins




 # Now, we can extract the outliers using boxplot.stats()$out function.  
 PublicWork_out <- boxplot.stats(Adata12$PublicWork)$out
 PublicWork_out
# 
# 
#  To count how many outliers in the population
  x9 <- data.frame(PublicWork_out)
 x9
  nrow(x9) # outliers numbers 
# 
# 
 #Now, we will remove the outliers from our dataset and store the clean result in Adata11
# 
   Adata13<- Adata12[-which(Adata12$PublicWork %in% PublicWork_out),] # here we removed outlier
# 
# 
 Adata13
#  
  hist(Adata13$PublicWork,
   xlab = "PublicWork",
   main = "Histogram of PublicWork",
   breaks = sqrt(nrow(Adata13))
) # set number of bins
  


# # No null values. 
 
# 
# # # Now we can calculate the max and the mean .. etc . 
 max(Adata13$PublicWork)
 mean(Adata13$PublicWork)
min (Adata13$PublicWork)
 



```





```{r step11, include=FALSE}

# same thing for PublicWork column. clean the data. 

# First, lets take a look at the histogram for the PublicWork to detect the outliers 



hist(Adata13$SelfEmployed,
  xlab = "SelfEmployed",
  main = "Histogram of SelfEmployed",
  breaks = sqrt(nrow(Adata13))
) # set number of bins




 # Now, we can extract the outliers using boxplot.stats()$out function.  
 SelfEmployed_out <- boxplot.stats(Adata13$SelfEmployed)$out
 SelfEmployed_out
# 
# 
#  To count how many outliers in the population
  x10 <- data.frame(SelfEmployed_out)
 x10
  nrow(x10) # outliers numbers 
# 
# 
 #Now, we will remove the outliers from our dataset and store the clean result in Adata11
# 
   Adata14<- Adata13[-which(Adata13$SelfEmployed %in% SelfEmployed_out),] # here we removed outlier
# 
# 
 Adata14
#  
  hist(Adata14$SelfEmployed,
   xlab = "SelfEmployed",
   main = "Histogram of SelfEmployed",
   breaks = sqrt(nrow(Adata14))
) # set number of bins
  


# # No null values. 
 
# 
# # # Now we can calculate the max and the mean .. etc . 
 max(Adata14$SelfEmployed)
 mean(Adata14$SelfEmployed)
min (Adata14$SelfEmployed)





```



**Final dataset **

As we can see bellow, after performing the cleansing process, we removed 1,147 record which include 2 NA values and 1,145 outliers. Now we have 2073 obs across 22 variables, the data shrink, however we have now more accurate dataset that will help us perform reliable tests as well as more accurate results.   



```{r final dataset, results='hold'}

# This is the final data set:
Adata9 = Adata14

str(Adata9)

#summary(Adata9)

#########################




```


**Final Dataset Summary **
```{r final summary, results='hold'}

summary(Adata9)


```


```{r SD, include=FALSE}


# Standard deviation for the columns: 

sd(Adata9$Income)
sd(Adata9$TotalPop)
sd(Adata9$Poverty)
sd(Adata9$ChildPoverty)
sd(Adata9$Employed)
sd(Adata9$Unemployment)
sd(Adata9$Production)
sd(Adata9$WorkAtHome)
sd(Adata9$PublicWork)
sd(Adata9$SelfEmployed)

```

### Income Prediction Test:

**How can we predict the average income in counties?**

First, we will use correlation matrix to see the relationships between different variables.

```{r income, results='hold'}

loadPkg("lattice")
loadPkg("ggplot2")
loadPkg("corrplot")




# To subset the data with numerical values only. 
data_num2 <- data.frame (subset(Adata9, select = -c(State,County,CensusId,Citizen,Service,Office,Professional, Construction, Production,WorkAtHome, TotalPop, Men, Women, FamilyWork)))


matrix2 = cor(data_num2) # get the correlation matrix between all numerical variables.
head(matrix2)
matrix2

corrplot(matrix2)


corrplot(matrix2, method = "number", type="upper")



```


**By using numerical variables only, we will build a linear model with 1 independent variable to predict the income.**  

We Choose a variable with strong correlation coefficient and here is our result:

1- Multiple R-squared value is equal to: 0.3 which means:
3% of the variability in the income is explained by the variability in Unemployment. 

2- The p-value is less than 0.05 which means the Unemployment is statistically significant predictor of income. 

3- The coefficient estimate value is equals to -1575.3 which means that there is 1575.3 decrease in the income when the Unemployment changed by 1. 

4- Coefficient-Standard Error= 52.9, so we can say that the income can vary by 52.9



```{r incomeCore, results='hold'}

cor.test(Adata9$Unemployment, Adata9$Income) -> cortest_1
cortest_1

fit_1 <- lm(Income ~ Unemployment, data = Adata9)
summary(fit_1)

xkabledply(fit_1, title = paste("Model :", format(formula(fit_1)) ) )

plot(fit_1)

```


**Next, we will add a second variable to the model.**  

We Choose a variable with stronger correlation coefficient and here is our result:

1- Multiple R-squared value is equal to: 0.675 which means:
67.5% of the variability in income is explained by the variability in unemployment and poverty.

2- The p-values of unemployment and poverty are less than 0.05 which means that they are statistically significant predictor of income. 

3- The coefficient estimate value of the unemployment is equals to -186.2 which means that there is 186.2 decrease in the income when the unemployment changed by 1. 

4- The coefficient estimate value of the poverty is equals to -1220.5 which means that there is 1220.5 decrease in the income when the poverty changed by 1. 

5- The Coefficient-Standard Error for the unemployment = 45.9, so we can say that the income can vary by 45.9

6- The Coefficient-Standard Error for the poverty = 25, so we can say that the income can vary by 25
```{r secondvarible, results='hold'}

fit_2 <- lm(Income ~ Unemployment+Poverty, data = Adata9)
summary(fit_2)




xkabledply(fit_2, title = paste("Model :", format(formula(fit_2)) ) )
xkablevif(fit_2)


```


**We will try one more time to add a third variable in our model.**


```{r thirdvarible, results='hold'}

fit_3 <- lm(Income ~ Unemployment+Poverty+ChildPoverty, data = Adata9)
summary(fit_3)



xkabledply(fit_3, title = paste("Model :", format(formula(fit_3)) ) )
xkablevif(fit_3)


```
**Confidence intervals of the coefficients.**  

```{r co3, results='hold'}

confint(fit_3)


```

**We used ANOVA to compare the three different models we found.**  

----------------------------------------------------------------------------------------------------------------

There is statistically significant difference in the mean of the income across the three models.

```{r anova11,results='hold'}

anovaRes_1 <- anova(fit_1,fit_2,fit_3) 
anovaRes_1
str(anovaRes_1)

xkabledply(anovaRes_1, title = "ANOVA comparison between the models") # to show it in a table

```
### Unemployment Prediction Test:

```{r Sagar}

#Sagar and Greesh put your code here


```




### Professional and Income Test:

```{r Alice}

# Alice put your code here. 

```


