---
title: "Project"
author: "Data lovers"
date: "10 oct 2021"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```



**importing the dataset as Adata, and making sure the data type for the variables are set properly for categorical variables.**

```{r step1}
Adata1 <- data.frame(read.csv("acs2015_county_data.csv",header = TRUE))

str(Adata1)

#summary(Adata1)

#  we don't need to change any variables type, all of them set properly.
# There are a lot of variables , we may drop some of them, like :
#Hispanic, White,	Black,	Native,	Asian	Pacific, IncomeErr,	IncomePerCap,	IncomePerCapErr


# Dropping columns as agreed.  

Adata <- data.frame (subset(Adata1, select = -c(Hispanic,White,Black, Native, Asian, Pacific,IncomeErr, IncomePerCap, IncomePerCapErr, Drive, Carpool, Transit, Walk, OtherTransp, MeanCommute)))

str(Adata)
summary(Adata)
  

```



***Cleaning the data. removing the outliers and the NA from the income. Saving the resulting dataframe as Adata2. **

```{r step2}

# After reading the data we need now to clean it to prepare for the analysis. 
# We started with "income", first we polte it to identify the outlier and then removing it. 
# Also , we removed the null values because it make error while do mathematics calculation. 
# Then we follow the same steps for the remaining columns that we will gonna use in this phase. 
# Currently our desired columns: TotalPop, Income, Poverty, ChildPoverty, Employed, Unemployment.  

################################################################################


# First, lets take a look at the histogram for the income to detect the outliers 



hist(Adata$Income,
  xlab = "Income",
  main = "Histogram of Income",
  breaks = sqrt(nrow(Adata))
) # set number of bins




# # Now, we can extract the outliers from the incom using boxplot.stats()$out function.  
 income_out <- boxplot.stats(Adata$Income)$out
 income_out


# To count how many outliers in the income
 x <- data.frame(income_out)
x 
 nrow(x) # outliers numbers for income


 #Now, we will remove the outliers from our dataset and store the clean result in Adata2

  Adata2<- Adata[-which(Adata$Income %in% income_out),] # here we removed income outlier


 Adata2
 
 hist(Adata2$Income,
  xlab = "Income",
  main = "Histogram of Income",
  breaks = sqrt(nrow(Adata2))
) # set number of bins

 
# #Adata
# Adata2 


 # Now we will remove the naull values from the incmoe to allow for math calculations. 
 
 my.null <- data.frame (subset(Adata2,subset= is.na(Adata2$Income)))# To indicate the null values of income. 
my.null


Adata3 <- data.frame (subset(Adata2, subset= CensusId!= 48301 )) # To remove the record that contains null value

# Now we can calculate the max, mean , min .. etc of the income
max(Adata3$Income)
mean(Adata3$Income)
min(Adata3$Income)



```




**Cleaning the data for population column. **

```{r}

# same thing for population column. clean the data. 

# First, lets take a look at the histogram for the population to detect the outliers 



hist(Adata3$TotalPop,
  xlab = "TotalPop",
  main = "Histogram of Population",
  breaks = sqrt(nrow(Adata3))
) # set number of bins




# # Now, we can extract the outliers from the population using boxplot.stats()$out function.  
 pop_out <- boxplot.stats(Adata3$TotalPop)$out
 pop_out


# To count how many outliers in the population
 x2 <- data.frame(pop_out)
x2 
 nrow(x2) # outliers numbers 


 #Now, we will remove the outliers from our dataset and store the clean result in Adata4

  Adata4<- Adata3[-which(Adata3$TotalPop %in% pop_out),] # here we removed  outlier


 Adata4
 
 hist(Adata4$TotalPop,
  xlab = "TotalPop",
  main = "Histogram of Population",
  breaks = sqrt(nrow(Adata4))
) # set number of bins

 
# #Adata
# Adata2 


 # Now we will remove the naull values from the population to allow for math calculations. 
 
 my.null2 <- data.frame (subset(Adata4,subset= is.na(Adata4$TotalPop)))# To indicate the null values of Population 
my.null2  # No null values. 


# Now we can calculate the max and the mean .. etc of the population. 
max(Adata4$TotalPop)
mean(Adata4$TotalPop)
min(Adata4$TotalPop)



```



**Cleaning the data for Poverty column. **
```{r}


# same thing for poverty column. clean the data. 

# First, lets take a look at the histogram for the poverty to detect the outliers 



hist(Adata4$Poverty,
  xlab = "Poverty",
  main = "Histogram of Poverty",
  breaks = sqrt(nrow(Adata4))
) # set number of bins




# # Now, we can extract the outliers from the population using boxplot.stats()$out function.  
 Poverty_out <- boxplot.stats(Adata4$Poverty)$out
 Poverty_out


# To count how many outliers in the population
 x3 <- data.frame(Poverty_out)
x3 
 nrow(x3) # outliers numbers 


 #Now, we will remove the outliers from our dataset and store the clean result in Adata5

  Adata5<- Adata4[-which(Adata4$Poverty %in% Poverty_out),] # here we removed  outlier


 Adata5
 
 hist(Adata5$Poverty,
  xlab = "Poverty",
  main = "Histogram of Poverty",
  breaks = sqrt(nrow(Adata5))
) # set number of bins

 
 



 # No null values. 
 

# Now we can calculate the max and the mean .. etc of the population. 
max(Adata4$Poverty)
mean(Adata4$Poverty)
min (Adata4$Poverty)






```


**Cleaning the data for Child Poverty column. **

```{r}


# same thing for Child Poverty column. clean the data. 

# First, lets take a look at the histogram for the Child Poverty to detect the outliers 



hist(Adata5$ChildPoverty,
  xlab = "ChildPoverty",
  main = "Histogram of ChildPoverty",
  breaks = sqrt(nrow(Adata5))
) # set number of bins




# # Now, we can extract the outliers using boxplot.stats()$out function.  
 ChildPoverty_out <- boxplot.stats(Adata5$ChildPoverty)$out
 ChildPoverty_out


# To count how many outliers in the population
 x4 <- data.frame(ChildPoverty_out)
x4
 nrow(x4) # outliers numbers 


 #Now, we will remove the outliers from our dataset and store the clean result in Adata6

  Adata6<- Adata5[-which(Adata5$ChildPoverty %in% ChildPoverty_out),] # here we removed  outlier


 Adata6
 
 hist(Adata6$ChildPoverty,
  xlab = "ChildPoverty",
  main = "Histogram of Child Poverty",
  breaks = sqrt(nrow(Adata6))
) # set number of bins

 
 

 
 my.null3 <- data.frame (subset(Adata6,subset= is.na(Adata6$ChildPoverty)))# To indicate the null values
my.null3  
 

 Adata7 <- data.frame (subset(Adata6, subset= CensusId!= 15005 )) # To remove the record that contains null value



# Now we can calculate the max and the mean .. etc of the population. 
max(Adata7$ChildPoverty)
mean(Adata7$ChildPoverty)
min (Adata7$ChildPoverty)






```


**Cleaning the data for Employed column  ** 

```{r}

# same thing for Employed column. clean the data. 

# First, lets take a look at the histogram for the Employed to detect the outliers 



hist(Adata7$Employed,
  xlab = "Employed",
  main = "Histogram of Employed",
  breaks = sqrt(nrow(Adata7))
) # set number of bins




# # Now, we can extract the outliers using boxplot.stats()$out function.  
 Employed_out <- boxplot.stats(Adata7$Employed)$out
 Employed_out


# To count how many outliers in the population
 x5 <- data.frame(Employed_out)
x5
 nrow(x5) # outliers numbers 


 #Now, we will remove the outliers from our dataset and store the clean result in Adata8

  Adata8<- Adata7[-which(Adata7$Employed %in% Employed_out),] # here we removed outlier


 Adata8
 
 hist(Adata8$Employed,
  xlab = "Employed",
  main = "Histogram of Employed",
  breaks = sqrt(nrow(Adata8))
) # set number of bins

 
 
# No null values. 
 

# # Now we can calculate the max and the mean .. etc of the population. 
max(Adata8$Employed)
mean(Adata8$Employed)
min (Adata8$Employed)



```


**Cleaning the data for Unemployment column  ** 


```{r}


# same thing for Unemployment column. clean the data. 

# First, lets take a look at the histogram for the Unemployment to detect the outliers 



hist(Adata8$Unemployment,
  xlab = "Unemployment",
  main = "Histogram of Unemployment",
  breaks = sqrt(nrow(Adata8))
) # set number of bins




 # Now, we can extract the outliers using boxplot.stats()$out function.  
 Unemployment_out <- boxplot.stats(Adata8$Unemployment)$out
 Unemployment_out
# 
# 
#  To count how many outliers in the population
  x6 <- data.frame(Unemployment_out)
 x6
  nrow(x6) # outliers numbers 
# 
# 
 #Now, we will remove the outliers from our dataset and store the clean result in Adata9
# 
   Adata9<- Adata8[-which(Adata8$Unemploymen %in% Unemployment_out),] # here we removed outlier
# 
# 
 Adata9
#  
  hist(Adata9$Unemployment,
   xlab = "Unemployment",
   main = "Histogram of Unemployment",
   breaks = sqrt(nrow(Adata9))
) # set number of bins
  


# # No null values. 
 
# 
# # # Now we can calculate the max and the mean .. etc . 
 max(Adata9$Unemployment)
 mean(Adata9$Unemployment)
min (Adata9$Unemployment)
 



```





**The final data set and calculating SD for our columns to check for data spread.** 


```{r}

# This is the final data set:
Adata9 

str(Adata9)

summary(Adata9)

#########################

# Standard deviation for the columns: 

sd(Adata9$Income)
sd(Adata9$TotalPop)
sd(Adata9$Poverty)
sd(Adata9$ChildPoverty)
sd(Adata9$Employed)
sd(Adata9$Unemployment)
```



**Normality tests  ** 
**The following graphs are the normality test for our desired columns.** 

**From these graphs, we noticed that the columns that use rates instead of total number are normally distributed like: income, poverty,child poverty, unemployment. While the columns that use total numbers are not normally distributed like: Total population and Employed columns.  ** 


```{r}


library(ggplot2)



#This code enable us to use qqplot.data function (to draw a line)

qqplot.data <- function (vec) # argument: vector of numbers
{
  y <- quantile(vec[!is.na(vec)], c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  int <- y[1L] - slope * x[1L]
  d <- data.frame(resids = vec)
  ggplot(d, aes(sample = resids)) + stat_qq() + geom_abline(slope = slope, intercept = int)
  
}

# Histogram that use ggplot: 

# 1 Graph for income distribution:
g1 <- qqplot.data(Adata9$Income)+labs(title = "Income Distribution")
g1

# 2 Graph for population distribution
g2 <- qqplot.data(Adata9$TotalPop)+labs(title = "Population Distribution")
g2


# 3 Graph for Poverty distribution
g3 <- qqplot.data(Adata9$Poverty)+labs(title = "Poverty Distribution")
g3


# 4 Graph for child poverty distribution
g4 <- qqplot.data(Adata9$ChildPoverty)+labs(title = "Child Poverty Distribution")
g4


# 5 Graph for Employed distribution
g5 <- qqplot.data(Adata9$Employed)+labs(title = "Employed Distribution")
g5


# 6 Graph for unemployment distribution
g5 <- qqplot.data(Adata9$Unemployment)+labs(title = "Unemployment Distribution")
g5

###########################################



# Additional graphs for two variables only.  QQ-plot that use qqnorm

# 1 Graph for income
qqnorm(Adata9$Income,main = "Q-Q plot. Income Distribution") 

#2 Graph for population
 qqnorm(Adata9$TotalPop,main = "Q-Q plot. Population Distribution")

 
#3 Graph for Poverty
 qqnorm(Adata9$Poverty,main = "Q-Q plot. Poverty Distribution")

 
#4 Graph for child poverty
 qqnorm(Adata9$ChildPoverty,main = "Q-Q plot. Child Poverty Distribution")

 #4 Graph for Employed
 qqnorm(Adata9$Employed,main = "Q-Q plot. Employed Distribution")
 
#5 Graph for unemployment
 qqnorm(Adata9$Unemployment,main = "Q-Q plot. Unemlpyment Distribution")
 



```












```{r}

# Ignore this chunk. 

#I just was trying to use leaflet to create maps. 

# install.packages("httpuv")
# 
# library(leaflet)
# 
# 
# water_locations_map <- leaflet(Adata3$Income) %>%
#   addTiles() %>%
#   addCircleMarkers(lng = ~location.longitude,
#                    lat = ~location.latitude)


# m <- leaflet() %>%
#   addTiles() %>%  # Add default OpenStreetMap map tiles
#   addMarkers(lng=174.768, lat=-36.852, popup="Adata3$Income")
# m  # Print the map




# avg = mean(Adata3$Income[!Adata3$Income])
# #avg 



# Adata_test=Adata3 
# 
# #Adata_test
# 
# Adata_test[which(Adata$Income %in% income_out),"Income"]= 0 # to replace the outliers values with 0 
# 
# Adata_test
#  
#  #income_out
#  
#  #df$Variable1[df$outlier] = avg
# 
# 
# 
# #Adata_test[which(Adata$Income %in% income_out),] = 4

```


### Q

**. **


```{r}


 
```


### Q

**? **

```{r, results='hold'}


```



### Q

**W **

```{r, results='hold'}


```



### Question 7

**f **

```{r Q7, results='hold'}



```



### Q

**.**

```{r Q8, results='hold'}



```





